---
title: "WK3 - In-Class Exercises"
author: "Jingyuan Yang - George Mason University, School of Business"
output:
  pdf_document: 
    latex_engine: xelatex
---

1 - Use the heart dataset and answer the following questions
Create a new dataset called heart_gender and use the split-apply-combine principle to group by gender, and ThalliumStresstest, and provide their count, avg age,  min restbp, max restbp, and avg maxhr

- Load the required libraries
```{r}
# Install and then load the library readxl and Tidyverse
library(????)
library(??)
# Load the content from the excel file into the data frame
heart <-  ??("???")

```

```{r}
heart_gender <- ?? %>%
  ??(Gender, ThalliumStressTest) %>%
  ??(gender_count = ??,
            avg_age = ??(Age),
            min_restbp =??(RestBP),
            max_restbp = ??(RestBP),
            avg_maxhr = ??(MaxHR))
heart_gender
```

2 - Create a new dataset - heart_age using the heart dataset. heart_age should have a new variable (AgeCategory), which combines the age and gender to provide the following categories (Use case_when)
Men30+,Women30+,MenUnder30, and WomenUnder30

```{r}
heart_age <- ?? %>%
  ??(AgeCategory = 
                  ??(
                       Age >= 30 ?? Gender =="??" ?? 'Men30+',
                       Age ?? 30 & Gender =="??" ~ 'Women30+',
                       Age ?? 30 & Gender =="male" ~ '??',
                       Age ?? 30 & Gender =="??" ~ 'WomenUnder30')
                  )
heart_age
```

3 - Use the heart_age dataset created above.
Create a RestBP_category variable that bins values in 10 increments
Form groups by AgeCategory and RestBP_category
Count the number of patients that fall into each combination

```{r}
?? %>% 
  ??(RestBP_category = ??(RestBP, ?? = 10, ?? = 0)) %>% 
  group_by(??,AgeCategory) %>% 
  summarize(Patients = ??)
```