---
title: "WK4 - In-Class Exercises"
author: "jiwon mok - George Mason University, School of Business"
output:
  pdf_document: 
    latex_engine: xelatex
---

q 1 - Use the daily show data which is also available in the package fivethirtyeight. But for now, use the data available in the link. Convert the existing variable to date variable. 
Also find the number of shows that were done on Tuesday and Friday


```{r}
library(readr)
library(dplyr)
library(stringr)

daily_show <- read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/daily-show-guests/daily_show_guests.csv',col_names = TRUE,show_col_types = FALSE )
show(daily_show)
```

```{r}
library(lubridate)
#New Variable with the day of the week
daily_show %>%
  mutate(Show = mdy(Show)) %>%
  mutate(Show_day = wday(Show, label=TRUE))%>%
  group_by(Show_day) %>% 
          summarise(count_wkday = n())
```

q2 Use the world cup data available in library(faraway) and use the data(worldcup)
How many players have the name with characters “ab” together
Find all players whose name begins with “S”.

```{r}
# Load the dataset
library(faraway)
data(worldcup)
#The dataset does not load the name, so create a new variable with player name
worldcup <- worldcup %>%
  mutate(player_name = rownames(worldcup))

```

```{r}
# Find all the players whose name has the character ab. 
# Hint - In this case, the characters could be at the beginning, between or end. This will also determine if you need to use lower case of upper case characters
worldcup %>%
  select(player_name) %>%
  filter(c(player_name, regex("ab", ignore_case = TRUE)))

```

```{r}
#Find all the players whose name starts with S
worldcup %>%
  select(player_name) %>%
  filter(??(player_name, "??"))
```



# Data Decisions

q3 Use the following vector
numeric_data <- c(7, 8, 9.5, 5.6, 11,1,-1,-22)
Use the ifelse() function to check if value is greater than or equal to 3 or less than 3
Also implement the if else logic to replace the 3rd element in the vector by 5, if the mean value is greater than 2, else replace the 4th vector by 3. 


- Load the required libraries
```{r}
# Part A
# Install and then load the Tidyverse
library(tidyverse)
# Load the content from the excel file into the data frame
numeric_data <- c(7, 8, 9.5, 5.6, 11,1,-1,-22)

# Check if value is greater than or equal to 3 or less than 3
less_3_check <- ifelse(??, "Less than 3", "3 or greater")

```

```{r}
# Part B
# Implement logic

# Assign the mean
numeric_data_mean <- numeric_data %>% ??
# Implement ifelse logic by using subsetting. 
ifelse(??, ??, ??)
numeric_data
```
q4 Use the map_dbl() (dbl is numeric) function to create a numeric vector that contains the column averages of all the variables in the built-in data frame, mtcars.
```{r}
# Use the library purrr
library(purrr)
# map_dbl with mean
mean_mtcars <- mtcars %>% ???
mean_mtcars
```


